---
title: "Microbiome analysis - data prep"
author: "Maureen A. Carey"
date: "09/01/2020"
output: html_document
---

Metadata variables: Metadata variables included age-related information (month/year of birth, age in days), socio-economic status (income, perceived food deficit), time of collection (month, year, amount of rainfall that month), the child’s home conditions (presence of an indoor drain, drinking water source, water treatment method, number of children under the age of 5 years), coinfection status (positive qPCR for Giardia spp. or E. histolytica), repeat infection v. primary infection, and growth status (z-scores for weight-for-age, height-for-age, weight-for-height, and body mass index for age). These are known to influence the development of diarrhea broadly or the acquisition of Cryptosporidium (Korpe et al. 2016; Steiner et al. 2018; Mondal et al. 2012)

Code for reference:
/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/open_and_save_all_data.Rmd to generate /Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData

The above code processes the private patient metadata and so cannot be shared at this point.

# Load data and packages

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/")

# load packages
library(phyloseq); library(reshape2); library(tibble)
library(rmarkdown); library(dplyr); library(tidyr)

# set constants
SEED_INT = 10; set.seed(SEED_INT); options(scipen=10000)
colors = c("#1F78B4","#FF7F00","#666666", 'light gray')
names(colors) = c('diarrheal','subclinical','no crypto','both')
# blue = "#1F78B4" orange = "#FF7F00" gray = "#666666"

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")

# reorganizing DOB files
DOB_df_with_study = DOB_df %>% filter(CHILDID %in% burden_mirpur$SID)
DOB_df = DOB_df %>% filter(CHILDID %in% burden_mirpur$SID) %>% select(-study)
left_study_ids = left_study_ids[left_study_ids>=2000 & left_study_ids <3406] # all files already have these kids removed

# subset for the cohort that is used for microbiome study
all_got_crypto = all_got_crypto %>% filter(CHILDID %in% burden_mirpur$SID)
all_kids = all_kids %>% filter(CHILDID %in% burden_mirpur$SID)
all_pathogens = all_pathogens %>% filter(CHILDID %in% burden_mirpur$SID)
burden_all_path = burden_all_path %>% filter(CHILDID %in% burden_mirpur$SID)
burden_crypto = burden_crypto %>% filter(CHILDID %in% burden_mirpur$SID)
burden_sex = burden_sex %>% filter(SID %in% burden_mirpur$SID)
children_without_diarrhea = children_without_diarrhea %>% filter(CHILDID %in% burden_mirpur$SID)
everything_df = everything_df %>% filter(CHILDID %in% burden_mirpur$SID)
first_infection = first_infection %>% filter(CHILDID %in% burden_mirpur$SID) # remaining kids never had a parasite
inflam_burden = inflam_burden %>% filter(CHILDID %in% burden_mirpur$SID)
total_abx_c = total_abx_c %>% filter(CHILDID %in% burden_mirpur$SID)        # remaining kids never took ABx
total_abx_c_with_diarr = total_abx_c_with_diarr %>% filter(CHILDID %in% burden_mirpur$SID)
total_diarrhea = total_diarrhea  %>% filter(CHILDID %in% burden_mirpur$SID) # remaining kids never were sick
crypto_cts = burden_all_path %>% filter(crypto == 1) %>%
  mutate(date_of_collection = as.Date(gsub('.{8}$', '', as.character(DOC)), "%m/%d/%Y")) %>%
  select(CHILDID, date_of_collection, CRYCTRT, type) %>% filter(CHILDID %in% burden_mirpur$SID)
all_cause_diar = total_diarrhea %>% group_by(CHILDID,study) %>% summarize(episodes = n()) %>% ungroup()

rm(list= ls()[! (ls() %in% c('left_study_ids','SEED_INT','colors','DOB_df',"burden_mirpur",'everything_df','total_diarrhea','burden_sex','crypto_cts'))])

```

# Generate supplemental file 1B to get read counts

```{r}

unadjusted_files_r1 = read.table("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/starting_seqs_run1.csv", 
                                 sep = ",", header = T, stringsAsFactors = F)
unadjusted_files_r2 = read.table("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/starting_seqs_run2.csv", 
                                 sep = ",", header = T, stringsAsFactors = F)
unadjusted_files_r1$X = NULL; unadjusted_files_r2$X = NULL
run1 = read.table("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/track_reads_run1.csv", 
                  sep = ",", header = T, stringsAsFactors = F); colnames(run1)[colnames(run1) == 'X'] = 'sample'
run2 = read.table("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/track_reads_run2.csv", 
                  sep = ",", header = T, stringsAsFactors = F); colnames(run2)[colnames(run2) == 'X'] = 'sample'
df1 = merge(run1, unadjusted_files_r1, by = "sample", all = T)
df2 = merge(run2, unadjusted_files_r2, by = "sample", all = T)
df = rbind(df1, df2); rm(df1, df2, run1, run2, unadjusted_files_r1, unadjusted_files_r2)

df_plot = melt(df %>% filter(!is.na(input))) %>% filter(variable %in% c("forward","input","filtered","nonchim"))
df_plot$variable = plyr::revalue(df_plot$variable, c("forward"="raw reads","input"="after adapter/\nprimer trimming", "filtered"="after quality\nfiltration", "nonchim"="after merging forward &\nreverse reads"))
df_plot$variable = factor(df_plot$variable, levels = c("raw reads","after adapter/\nprimer trimming", "after quality\nfiltration", "after merging forward &\nreverse reads"), ordered = T)
df_mean = df_plot %>% group_by(variable) %>% summarize(mean = mean(value), median = median(value))

df_summary = df
read_count_stats_df = df_mean
read_counts_df = df_plot

```

## Reading the phyloseq object

Read in the phyloseq object generated via dada2. Remove any mitochondrial or chloroplast reads.

```{r}

ps = readRDS("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/ps_2020_08_12.rds")

# Quality control - remove mitochondrial and chloroplast reads
og_taxa = data.frame(tax_table(ps))
# first remove non-bacteria
keep = rownames(og_taxa[og_taxa$Kingdom %in% c("Bacteria","Archaea"),])
ps = prune_taxa(keep,ps)
keep = rownames(og_taxa[og_taxa$Family != "Mitochondria",])
ps = prune_taxa(keep,ps)
keep = rownames(og_taxa[og_taxa$Class != "Chloroplast",])
ps = prune_taxa(keep,ps)

rm(df, keep, og_taxa)

```

Prep plot positive controls

```{r}

pos_controls1 = subset_samples(ps, startsWith(rownames(sample_data(ps)), 'pos') & endsWith(rownames(sample_data(ps)), 'run1'))
pos_controls2 = subset_samples(ps, startsWith(rownames(sample_data(ps)), 'pos') & endsWith(rownames(sample_data(ps)), 'run2'))
rare_min = 10000
pos_controls1= rarefy_even_depth(pos_controls1,sample.size=rare_min,rngseed=SEED_INT)
pos_controls2= rarefy_even_depth(pos_controls2,sample.size=rare_min,rngseed=SEED_INT)

# run 1
pos_controls = pos_controls1; top_quantity = 16
# Get the most abundant ASVs across the entire dataset
most_abundant_names = paste(1:top_quantity,
                            tax_table(pos_controls)[names(sort(taxa_sums(pos_controls),TRUE)[1:top_quantity]),"Genus"],"")
most_abundant_genus = tax_table(pos_controls)[names(sort(taxa_sums(pos_controls), TRUE)[1:top_quantity]),"Genus"]
most_abundant = names(sort(taxa_sums(pos_controls), TRUE)[1:top_quantity])
ps_for_abundplot = prune_taxa(most_abundant,pos_controls)
readsumsdf = data.frame(nreads = sort(taxa_sums(ps_for_abundplot), TRUE), sorted = 1:ntaxa(ps_for_abundplot),  type = "OTUs")
readsumsdf$names = factor(most_abundant_names,levels=most_abundant_names)
readsumsdf$genus = factor(most_abundant_genus)
rownames(readsumsdf) = NULL

plot_run1 = readsumsdf
rm(pos_controls, top_quantity,most_abundant_names, most_abundant, most_abundant_genus,ps_for_abundplot,
   readsumsdf)

# run 2
pos_controls = pos_controls2; top_quantity = 15
# Get the most abundant ASVs across the entire dataset
most_abundant_names = paste(1:top_quantity,
                            tax_table(pos_controls)[names(sort(taxa_sums(pos_controls),TRUE)[1:top_quantity]),"Genus"],"")
most_abundant_genus = tax_table(pos_controls)[names(sort(taxa_sums(pos_controls), TRUE)[1:top_quantity]),"Genus"]
most_abundant = names(sort(taxa_sums(pos_controls), TRUE)[1:top_quantity])
ps_for_abundplot = prune_taxa(most_abundant,pos_controls)
readsumsdf = data.frame(nreads = sort(taxa_sums(ps_for_abundplot), TRUE), sorted = 1:ntaxa(ps_for_abundplot),  type = "OTUs")
readsumsdf$names = factor(most_abundant_names,levels=most_abundant_names)
readsumsdf$genus = factor(most_abundant_genus)
rownames(readsumsdf) = NULL
plot_run2 = readsumsdf

rm(pos_controls, top_quantity, rare_min,most_abundant_names, most_abundant, most_abundant_genus,ps_for_abundplot,
   readsumsdf)

```

## Get only Mirpur samples.

```{r}

ps = subset_samples(ps, Site == 1) 
ps = subset_samples(ps, !startsWith(rownames(sample_data(ps)), 'pos'))

# remove if error in any reporting re Crypto infection
keep1 = ifelse(sample_data(ps)$Crypto..Any. == '1' & as.numeric(sample_data(ps)$Clinical.type) < 3, 0, 1)
ps = prune_samples(as.logical(keep1),ps)
keep2 = ifelse(sample_data(ps)$Crypto..Any. == '999' & as.numeric(sample_data(ps)$Clinical.type) > 3, 0, 1)
ps = prune_samples(as.logical(keep2),ps)
keep3 = ifelse(sample_data(ps)$Crypto..Any. == '888' & as.numeric(sample_data(ps)$Clinical.type) > 3, 0, 1)
ps = prune_samples(as.logical(keep3),ps)

rm(keep1, keep2, keep3)

```

Convert sample data to factors and convert dates to month only for easier handling.

```{r}

# convert some sample data to factors
tempsampledata = sample_data(ps)
tempsampledata$Site = NULL
tempsampledata$Clinical.type = as.factor(tempsampledata$Clinical.type)
tempsampledata$Number.of.crypto.positive.events = as.factor(tempsampledata$Number.of.crypto.positive.events)
# Clinical types
# 12=diarrhea; 10=Clinic Visit; 1 =pre-clinic visit positive; 2=pre-diarrhea case positive
tempsampledata$symptomatic = tempsampledata$Clinical.type %in% c(2,12)
tempsampledata$pre_detection = tempsampledata$Clinical.type %in% c(1,2)

# fix dates
tempsampledata$Date.of.collection = as.Date(as.character(tempsampledata$Date.of.collection),format = "%d-%B-%y")
tempsampledata$year.of.collection = as.factor(as.numeric(format(tempsampledata$Date.of.collection,'%Y')))
tempsampledata$month.of.collection = as.factor(as.numeric(format(tempsampledata$Date.of.collection,'%m')))
tempsampledata$Date.of.birth = as.Date(as.character(tempsampledata$Date.of.birth),format = "%d-%B-%y")
tempsampledata$year.of.birth = as.factor(as.numeric(format(tempsampledata$Date.of.birth,'%Y')))
tempsampledata$month.of.birth = as.factor(as.numeric(format(tempsampledata$Date.of.birth,'%m')))

# ONE PROBLEM => STILL A PROBLEM
sum(tempsampledata$Age.in.days != tempsampledata$Date.of.collection-tempsampledata$Date.of.birth)

# Add rainfall data
rainfall = read.table("~/local_documents/work/crypto_microbiome/data_cannot_be_shared/Rainfall.csv",header=TRUE, quote="",sep='\t') %>% filter(Site == 1) %>% select(-Site)
rainfall$month.of.collection = as.factor(rainfall$month)
rainfall$year.of.collection = as.factor(paste('20',rainfall$year, sep = ''))
rainfall$rainfall = rainfall$Rainfall..mm.
rainfall$Rainfall..mm. = NULL; rainfall$month = NULL; rainfall$year = NULL
tempsampledata = merge(tempsampledata,rainfall,by=c("month.of.collection","year.of.collection"), all.x = T)

# convert all "Any assays to factors"
any_colnames <- names(tempsampledata)[grep("Any",names(tempsampledata))]
tempsampledata[,any_colnames] <- lapply(tempsampledata[,any_colnames], factor)
any_colnames <- names(tempsampledata)[grep("any",names(tempsampledata))]
tempsampledata[,any_colnames] <- lapply(tempsampledata[,any_colnames], factor)

# Correct for sample name-dropping that occurs when merging
tempsampledata = sample_data(tempsampledata)
sample_names(tempsampledata) = tempsampledata$sample.id

# save
sample_data(ps) = tempsampledata
rm(rainfall,any_colnames, tempsampledata)

# remove children who were lost to follow up
ps = subset_samples(ps, !(Child.ID %in% left_study_ids))

```

## Adding genotype, pan-crypto Ct data, and species-specific Ct data

These data were obtained after the initial processing of the sequence data, so we will add them to the metadata now.

```{r}

tempsampledata = as(sample_data(ps),"data.frame")

# Add genotype information for all samples
tempsampledata = merge(tempsampledata %>% tibble::rownames_to_column('use_as_rows'), 
                       everything_df %>% select(-DOB,-study,-birth_order,-WAZ_2yr,-HAZ_2yr,-WHZ_2yr), 
                       by.x = "Child.ID", by.y = "CHILDID", all.x = T)
tempsampledata$drain = as.factor(tempsampledata$drain)
tempsampledata$food_deficit = as.factor(tempsampledata$food_deficit)
tempsampledata$toilet = as.factor(tempsampledata$toilet)
tempsampledata$drinking_water_source = as.factor(tempsampledata$drinking_water_source)

# Add crypto Cts
tempsampledata$Child.ID = as.character(tempsampledata$Child.ID)
tempsampledata = merge(tempsampledata, crypto_cts %>% select(-type), by.x = c("Child.ID","Date.of.collection"), by.y = c("CHILDID","date_of_collection"), all.x = T)
tempsampledata[tempsampledata$pre_detection == T,"CRYCTRT"] = 45 # if predetection, Crypto DNA was not assessed or detected

# get other Ct values (targeted qPCRs)
all_ct = read.table("~/local_documents/work/crypto_microbiome/data_cannot_be_shared/Cq_values_all.txt",header=TRUE, sep='\t',comment.char = "",stringsAsFactors = F, na.strings = c("#N/A"))
all_ct[is.na(all_ct)] = 45 # replace #NA and real NAs with 0
all_ct[all_ct == 0] = 45 # replace #NA and real NAs with 0
tempsampledata = tempsampledata %>% mutate(sample_id_mod = substr(sample.id,1,nchar(sample.id)-5))
tempsampledata = merge(tempsampledata, all_ct, by.x = "sample_id_mod", by.y = "sample.id", all.x = T)
rownames(tempsampledata) = tempsampledata$use_as_rows; tempsampledata$use_as_rows = NULL
sample_data(ps) = tempsampledata

df = all_ct %>% mutate(double_pos = ifelse(LIB13.C..hominis<45 & C..mel<45,1,0)) %>%
  melt(id.vars = c("Pan_crypto",'sample.id','double_pos')) %>%
  mutate(infected = ifelse(double_pos,'double positive',ifelse(Pan_crypto<45,'single',ifelse(value<45,'single','not')))) %>%
  dcast(...~variable, value = 'value') %>%
  mutate(infected_use = ifelse(LIB13.C..hominis<45 & C..mel<45,'double positive',
                               ifelse(LIB13.C..hominis<45, 'C. hominis',
                                      ifelse(C..mel<45, 'C. meleagridis','not')))) %>%
  subset(infected != "not") %>% 
  melt(id.vars = c("Pan_crypto",'sample.id','double_pos','infected','infected_use')) 

df = rbind(rbind(df[df$infected == 'single' & df$variable == 'LIB13.C..hominis' & df$infected_use == 'C. hominis' & df$value > 0,],
           df[df$infected == 'single' & df$variable == 'C..mel' & df$infected_use == 'C. meleagridis' & df$value > 0,]),
           df[df$infected == 'double positive' ,])

ct_df = df

rm(df,tempsampledata,all_ct)

```

### Redo figure with Cts
```{r}

crypto_symptoms = merge(crypto_cts, DOB_df %>% select(-DOB), by = "CHILDID", all = T) %>% #all x only
  select(-date_of_collection) %>%
  mutate(type_use = ifelse(type == 'subclinical', 1, ifelse(type == 'diarrheal', 2, 0))) %>% select(-type) %>%
  mutate(type = ifelse(is.na(type_use), 0, type_use),
         site = ifelse(CHILDID %in% burden_mirpur$SID, 'Mirpur','Mirzapur')) %>% select(-type_use) %>%
  group_by(CHILDID, type, site) %>% 
  summarize(min_ct = ifelse(sum(is.na(CRYCTRT)) == length(CRYCTRT),45,
                            min(CRYCTRT,na.rm = T))) %>% ungroup() %>% 
  group_by(CHILDID, site) %>% summarize(val = mean(type), min_ct_overall = min(min_ct,na.rm = T)) %>% 
  mutate(TYPE = ifelse(val == 0, 'never infected',
                       ifelse(val == 2, 'diarrheal', 
                              ifelse(val == 1, 'subclinical',
                                     'both')))) %>% select(-val)

temp_df = merge(total_diarrhea, crypto_symptoms, by = "CHILDID", all = T)

# get kids that had no diarrhea
no_diarr = temp_df[is.na(temp_df$EPINUM),] %>% mutate(n_diar = 0) %>% select(n_diar, TYPE, CHILDID, min_ct_overall)

# summarize for kids who had diarrhea
yes_diarr = temp_df %>% filter(!is.na(EPINUM)) %>% group_by(CHILDID) %>% mutate(n_diar = n()) %>% select(n_diar, TYPE, CHILDID, min_ct_overall) %>% dplyr::slice(1) %>% ungroup()

# join 
df_plot = rbind(no_diarr, yes_diarr) 
df_plot$TYPE <- factor(df_plot$TYPE, levels = c("never infected","subclinical","diarrheal","both" ))
levels(df_plot$TYPE)[levels(df_plot$TYPE) == 'never infected'] = 'no infection\ndetected'
ct_updated_df = df_plot

rm(no_diarr, yes_diarr,temp_df,df_plot)

```

To correct for differences in sequencing depth across samples, we will subsample to a uniform depth of 10000 reads for each sample.

```{r message=FALSE}

# Rarify reads in all samples
rare_min = 10000
# remove samples with fewer counts than the minimum
ps_rare = prune_samples(sample_sums(ps) >=rare_min,ps) #8 removed
# remove samples with NA age
ps_rare = prune_samples(!is.na(sample_data(ps_rare)[,"Age.in.days"])[,1],ps_rare) # none removed
ps_rare = rarefy_even_depth(ps_rare,sample.size=rare_min,rngseed=SEED_INT) # OTUs no longer in the dataset will be removed
num_samples = length(sample_sums(ps_rare))

```

Add diversity to metadata.

```{r}

# adding Inverse Simplson to metadata 
sample_data(ps_rare)$InvSimp = as.numeric(unlist(estimate_richness(ps_rare,measures="InvSimpson")))

# adding Numebr of ASVs to metadata
richness = estimate_richness(ps_rare,measures="Observed")
colnames(richness) = c("NumASVs")

row.names(richness) = gsub("X","",row.names(richness)) # for some reason, estimate richness adds an "X" at the beginneing of row names.
asdf = as(sample_data(ps_rare),"data.frame")
asdf = merge(asdf,richness,by="row.names")
row.names(asdf) = asdf$Row.names
asdf = asdf[,!colnames(asdf) %in% c("Row.names")]
asdf$NumASVs = as.numeric(unlist(asdf$NumASVs))
sample_data(ps_rare) = asdf

rm(asdf, richness,num_samples)

```

## computation of change in each ASV between pre- and post- detection events

```{r}

# select only the pre/post detection samples, which will exclude any clinical type other than 1,2,10, or 12.
# also exclude repeat infections (only get first)
ps_rare_diff = subset_samples(ps_rare,sample_data(ps_rare)$Clinical.type %in% c(1,2,10,12))
tempsampledata = as(sample_data(ps_rare_diff),"data.frame")
new_data = tempsampledata
new_data$Clinical.type = as.numeric(as.character(new_data$Clinical.type))
    
# Add the sample ID for the previous surveilance sample and the age of the child at that time
new_data$previous_surv = NA; new_data$age_at_surv = NA
new_data$pre_InvSimp = NA; new_data$pre_NumASVs = NA
for (i in 1:nrow(new_data)) {
  # get TOD info
  kid = new_data$Child.ID[i]
  age_during_this_sample = new_data$Age.in.days[i] # in days
  clinical_type_of_this = new_data$Clinical.type[i]

  # make sure we're looking at the right group of pre_detection samples
  if (clinical_type_of_this == 10) { desired_clinical_type = 1
  } else if (clinical_type_of_this == 12) { desired_clinical_type = 2
  } else { desired_clinical_type = NA }
  one_kids_other_samples = new_data[new_data$Child.ID == kid,] %>% filter(Clinical.type == desired_clinical_type)

  # make sure pre_detection sample is within 30 days
  # if one option
  if ( nrow(one_kids_other_samples) == 1 ) {
    if (one_kids_other_samples$Age.in.days < age_during_this_sample & one_kids_other_samples$Age.in.days > age_during_this_sample - 42) {
      new_data$previous_surv[i] = one_kids_other_samples$sample.id
      new_data$age_at_surv[i] = one_kids_other_samples$Age.in.days
      new_data$pre_InvSimp[i] = one_kids_other_samples$InvSimp
      new_data$pre_NumASVs[i] = one_kids_other_samples$NumASVs }
  # if one option
    } else if ( nrow(one_kids_other_samples) > 1) {
      for (x in 1:nrow(one_kids_other_samples) ) {
        if (one_kids_other_samples$Age.in.days[x] < age_during_this_sample & one_kids_other_samples$Age.in.days[x] > age_during_this_sample - 42) {
          new_data$previous_surv[i] = one_kids_other_samples$sample.id[x]
          new_data$age_at_surv[i] = one_kids_other_samples$Age.in.days[x]
          new_data$pre_InvSimp[i] = one_kids_other_samples$InvSimp[x]
          new_data$pre_NumASVs[i] = one_kids_other_samples$NumASVs[x] }
        }
      } else { next }
}
rm(age_during_this_sample,one_kids_other_samples,desired_clinical_type,kid,i,x)

new_data$Clinical.type = as.factor(as.character(new_data$Clinical.type))
use_this = sample_data(new_data)
sample_names(use_this) = use_this$sample.id
sample_data(ps_rare_diff) = use_this

#omit if predetection sample doesn't have a paired detection sample
ps_rare_diff = (subset_samples(ps_rare_diff, (sample_data(ps_rare_diff)$sample.id %in% sample_data(ps_rare_diff)$previous_surv) | sample_data(ps_rare_diff)$pre_detection == F))
#omit if detection sample doesn't have a piared predetection sample
ps_rare_diff = (subset_samples(ps_rare_diff, !(is.na(sample_data(ps_rare_diff)$previous_surv)) | sample_data(ps_rare_diff)$pre_detection == T))

rm(new_data,tempsampledata,use_this, clinical_type_of_this)

for_stats = ps_rare_diff

```

# match samples and get CHANGE in microbiome
``` {r}

# get dataset where samples are matched
ps_rare_for_beta = ps_rare_diff

# calculate difference
temp_otu = otu_table(ps_rare_for_beta)
keep_these = c()
for (i in 1:nrow(sample_data(ps_rare_for_beta))) {
  time_of_detection_sample_id = rownames(sample_data(ps_rare_for_beta))[i]
  predetection_sample_id = sample_data(ps_rare_for_beta)$previous_surv[i]
  # make true if NA to remove from dataset
  if (is.na(time_of_detection_sample_id)) {sample_data(ps_rare_for_beta)$pre_detection[i] = T}
  if (is.na(predetection_sample_id)) {sample_data(ps_rare_for_beta)$pre_detection[i] = T}
  
  if (!sample_data(ps_rare_for_beta)$pre_detection[i]) {
    keep_these[i] = T
    temp_otu[rownames(temp_otu) == time_of_detection_sample_id,] = 
      temp_otu[rownames(temp_otu) == time_of_detection_sample_id,] - temp_otu[rownames(temp_otu) == predetection_sample_id,]
  }
  else { # we are looking at a predetection sample and aren't interested in it
    keep_these[i] = F }
}
# resave
otu_table(ps_rare_for_beta) = temp_otu

# remove predetection samples
ps_rare_for_beta = prune_samples(!sample_data(ps_rare_for_beta)$pre_detection,ps_rare_for_beta)

# get samples with fewer than 5 non-zero OTUs
nonzero_count = apply(otu_table(ps_rare_for_beta), 1, function(c)sum(c!= 0))
ps_rare_for_beta = prune_samples(nonzero_count > 5, ps_rare_for_beta)
change_in_microbiota_ps = ps_rare_for_beta
dis_df = dist(otu_table(ps_rare_for_beta))
pca_df = prcomp(dis_df)
axis_location = cbind(pca_df$rotation, sample_data(ps_rare_for_beta))

axis_location$symptomatic = as.factor(axis_location$symptomatic)
levels(axis_location$symptomatic)[levels(axis_location$symptomatic)=="FALSE"] = "subclinical"
levels(axis_location$symptomatic)[levels(axis_location$symptomatic)=="TRUE"] = "diarrheal"

```

# Save data for other scripts

```{r}

save(list = c("df_summary","read_count_stats_df", "read_counts_df","pos_controls1","pos_controls2","plot_run1", "plot_run2", "ct_df","ct_updated_df","ps_rare_diff","ps","axis_location","for_stats","ps_rare","change_in_microbiota_ps","burden_sex"), file = paste0(paste0("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_plot_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

save(list = c("for_stats","ps","ps_rare","ps_rare_diff","change_in_microbiota_ps"), file = paste0(paste0("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_" , format(Sys.time(), "%Y_%m_%d")),".RData")) 

```

# Make figures

```{r}

#rmarkdown::render("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/code/R_markdown/P1_analysis_prep_figures.Rmd")

```

