---
title: "Microbiome analysis - machine learning"
author: "Maureen A. Carey"
date: "09/01/2020"
output: html_document
---

# Load packages and data

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/")

# load packages
library(phyloseq); library(reshape2); library(tibble)
library(randomForest); library(rmarkdown); library(dplyr); library(tidyr); library(pROC)

# set constants
SEED_INT = 10; set.seed(SEED_INT); options(scipen=10000)
colors = c("#1F78B4","#FF7F00","#666666", 'light gray')
names(colors) = c('diarrheal','subclinical','no crypto','both')
# blue = "#1F78B4" orange = "#FF7F00" gray = "#666666"

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
total_diarrhea = total_diarrhea  %>% filter(CHILDID %in% burden_mirpur$SID) # remaining kids never were sick
all_cause_diar = total_diarrhea %>% group_by(CHILDID,study) %>% summarize(episodes = n()) %>% ungroup()
rm(list= ls()[! (ls() %in% c('all_cause_diar','SEED_INT','colors'))])
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_02.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P2output_2020_09_02.RData")

# make useful functions
randomForest_res = function(x,y,N) { 
  # x = variables used to predict class, y = classs labels, N = number imp variables you want
  # THIS ONLY WORKS IF YOU"RE CLASSIFYING 10 v. 12 (subclinical v. diarrheal)
  rf_res_object = randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.0001,100))

  extracted_importance = importance(rf_res_object,type=2) # type 2 = node impurity, more useful for auc
  extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
  topN = tail(extracted_importance,N)

  predictions = as.data.frame(predict(rf_res_object,type="prob"))
  predictions$predict = names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
  predictions$observed = y
  
  ROC_res = roc(ifelse(predictions$observed==10, 10, 12), as.numeric(predictions$`10`))
  AUC_res = auc(ROC_res)
  return(list(rf_res_object,ROC_res,AUC_res,topN))  
}

```

# Can we predict whether a sample was sub-clinical or diarrheal from predetection microbiota?

```{r}

compare <- c(1,2) # predetection -> +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)

x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
rf <- randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.00015,130))
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_PD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))   
x_for_PD = x
y_for_PD = y

extracted_importance = importance(rf,type=2) # type 2 = node impurity, more useful for auc
extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
top10_PD = tail(extracted_importance,10)

predictions = as.data.frame(predict(rf,type="prob"))
predictions$predict <- names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
predictions$observed <- y

roc_PD <- roc(ifelse(predictions$observed==1, 1, 2), as.numeric(predictions$`1`))
AUC_PD = auc(roc_PD)
AUC_PD

rm(x,y,rf,predictions,extracted_importance,ps_for_class)

```

# Can we predict whether a sample was sub-clinical or diarrheal from TOD microbiota?

```{r}

compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
x_for_TOD = x
y_for_TOD = y

rf = randomForest_res(x_for_TOD,y_for_TOD,10)
top10_TOD = rf[[4]]
roc_TOD = rf[[2]]
AUC_TOD = rf[[3]]
AUC_TOD

rf = rf[[1]]
# use this on rivanna for faster computing
save(list = c("rf","x","y"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_TOD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  
rm(x,y,rf,ps_for_class)

```

# Save data for other scripts

```{r}

save(list = c("top10_TOD","top10_PD"), file = paste0(paste0("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p1_" , format(Sys.time(), "%Y_%m_%d")),".RData")) 

```

## DATA NOT SHOWN: Can we predict whether a sample was a first or second infection from TOD microbiota?

```{r}

# time-of-detection
compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.numeric(as.character(sample_data(ps_for_class)$Number.of.crypto.positive.events))
y[y!=1] = 2 # merge third infections with second

temp_df = as(sample_data(ps_for_class),"data.frame") %>% 
  select(Number.of.crypto.positive.events, symptomatic) %>%
  mutate(n_crypto = ifelse(Number.of.crypto.positive.events == 1, '1', '2+')) %>% select(-Number.of.crypto.positive.events) %>%
  table() %>% as.data.frame() %>%
  mutate(symptomatic_use = ifelse(symptomatic == T, 'diarrheal','subclinical')) %>%
  select(-symptomatic) %>% dplyr::rename(symptomatic = symptomatic_use)
chi_df = temp_df

rf = randomForest(x,as.factor(y),importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.1,100))
rf
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_TOD_1stv2nd_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

extracted_importance = importance(rf,type=2) # type 2 = node impurity, more useful for auc
extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
topN_first_infection = tail(extracted_importance,10)

predictions = as.data.frame(predict(rf,type="prob"))
predictions$predict = names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
predictions$observed = y

ROC_first_infection = roc(ifelse(predictions$observed==1, 1, 2), as.numeric(predictions$`1`))
auc(ROC_first_infection) 

rm(x,y,rf,ps_for_class)

```

## DATA NOT SHOWN: Can we predict whether a sample was a first or second infection from PD microbiota?

```{r}

# predetection
compare <- c(1,2) # PD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.numeric(as.character(sample_data(ps_for_class)$Number.of.crypto.positive.events))
y[y!=1] = 2 # merge third infections with second
rf = randomForest(x,as.factor(y),importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.1,100))
rf
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_PD_1stv2nd_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

extracted_importance = importance(rf,type=2) # type 2 = node impurity, more useful for auc
extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
topN_first_infection_PD = tail(extracted_importance,10)

predictions = as.data.frame(predict(rf,type="prob"))
predictions$predict = names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
predictions$observed = y

ROC_first_infection = roc(ifelse(predictions$observed==1, 1, 2), as.numeric(predictions$`1`))
auc(ROC_first_infection) 

rm(x,y,rf,ps_for_class)

```

## DATA NOT SHOWN: Can we predict all cause diarrhea from TOD microbiota?

```{r}

compare <- c(10,12) # TOD +/- symptoms
ps_for_class2 <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
all_cause_diar_temp = all_cause_diar %>% filter(CHILDID %in% sample_data(ps_for_class2)$Child.ID)
all_cause_diar_temp = merge(all_cause_diar_temp, as(sample_data(ps_for_class2)[,1:2], "data.frame"), by.x = "CHILDID", by.y = "Child.ID", all.y = T) %>% select(-study,-CHILDID)
sample_data(ps_for_class2)$episodes = all_cause_diar_temp$episodes[order(match(all_cause_diar_temp$sample_id_mod, sample_data(ps_for_class2)$sample_id_mod))]
x <- cbind(otu_table(ps_for_class2), as(sample_data(ps_for_class2),"data.frame") %>% select(Age.in.days))
y <- sample_data(ps_for_class2)$episodes
## WHY IS THERE AN NA
y[is.na(y)] = 0

rf = randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.0001,100))
rf

save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_all_diar_fromTOD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

rm(x,y,rf,ps_for_class2)

```

## DATA NOT SHOWN: Can we predict all cause diarrhea from PD microbiota?

```{r}

compare <- c(1,2) # PD +/- symptoms
ps_for_class2 <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
all_cause_diar_temp = all_cause_diar %>% filter(CHILDID %in% sample_data(ps_for_class2)$Child.ID)
all_cause_diar_temp = merge(all_cause_diar_temp, as(sample_data(ps_for_class2)[,1:2], "data.frame"), by.x = "CHILDID", by.y = "Child.ID", all.y = T) %>% select(-study,-CHILDID)
sample_data(ps_for_class2)$episodes = all_cause_diar_temp$episodes[order(match(all_cause_diar_temp$sample_id_mod, sample_data(ps_for_class2)$sample_id_mod))]
x <- cbind(otu_table(ps_for_class2), as(sample_data(ps_for_class2),"data.frame") %>% select(Age.in.days))
y <- sample_data(ps_for_class2)$episodes
## WHY IS THERE AN NA
y[is.na(y)] = 0

rf = randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.0001,100))
rf

save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_all_diar_fromPD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

rm(x,y,rf,ps_for_class2)

```

## Can we predict whether a sample was sub-clinical or diarrheal from PD & TOD microbiota?
# not a great analysis because trees don't scale well to compare AUC

```{r}

compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
ps_for_class <- subset_samples(ps_for_class, !is.na(previous_surv))
ps_for_surv <- subset_samples(ps_rare_diff, sample.id %in% sample_data(ps_for_class)$previous_surv)
taxa_names(ps_for_surv) <- paste0("pre_", taxa_names(ps_for_surv))
time_of_detect_microbes = otu_table(ps_for_class)
surv_microbes = otu_table(ps_for_surv)
# reorder the surveilance to match the TOD
proper_order = sample_data(ps_for_class)$previous_surv
surv_microbes = surv_microbes[proper_order]

x <- cbind(cbind(time_of_detect_microbes,surv_microbes), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y = as.factor(sample_data(subset_samples(ps_for_class, Clinical.type %in% compare))$Clinical.type)
rf = randomForest_res(x,y,15)
rf[[1]]

top_15_for_all_microbiome = rf[[4]]
roc_allM = rf[[2]]
AUC_allM = rf[[3]]
AUC_allM
rm(rf, x,y,proper_order,ps_for_surv)

```

## Can we predict whether a sample was sub-clinical or diarrheal from PD & TOD microbiota AND metadata?
# not a good analysis because trees don't scale well to compare AUC

```{r}
 
meta_for_rf = as(sample_data(ps_for_class)[,c( "rainfall","drain","food_deficit",
                                                #"toilet", # too many NAs
                                                "drinking_water_source","income_mo",
                                                "water_treat","siblings_under_5",
                                                #"CREM_rs2148483_A","PRKCA_rs58296998_T", #too many NAs
                                                "month.of.birth","year.of.birth",
                                                "month.of.collection",
                                                "year.of.collection", "Any..Giardia",
                                                "Any..E.histolytica",
                                                "WAZ","WHZ","HAZ",#"MUACZ",
                                               "BAZ",
                                                "Number.of.crypto.positive.events",
                                                "Age.in.days")],"data.frame")
meta_for_rf$water_treat = as.factor(meta_for_rf$water_treat)

x = cbind(time_of_detect_microbes,meta_for_rf,surv_microbes)
y = as.factor(sample_data(subset_samples(ps_for_class, Clinical.type %in% compare))$Clinical.type)
rf = randomForest_res(x,y,20)
rf[[1]]

top15_without_abx = rf[[4]]
roc_allM_M = rf[[2]]
AUC_allM_M = rf[[3]]
AUC_allM_M

rm(rf,x,y,ps_for_class)

```

# Prep Abx data for integration into model

```{r}

for_metadata = subset_samples(ps_rare_diff, Clinical.type %in% c(10,12))
for_metadata = subset_samples(for_metadata, !is.na(previous_surv))
temp_df = as(sample_data(for_metadata)[,c("sample.id")],"data.frame") 
rownames(temp_df) = temp_df$sample.id
temp_df2 = as(sample_data(for_metadata)[,c("Child.ID","Date.of.collection","sample.id")],"data.frame")
# kids with zero ABx don't have their Run# in these dataframes - cut those kids and then add after
keep_for_rf_all_abx2 = merge(keep_for_rf_all_abx %>% filter(!is.na(sample.id)), temp_df, 
                            by = 'sample.id', all.y = T) %>% select(sample.id,n_abx_doses)
keep_for_rf_all_abx2$n_abx_doses[is.na(keep_for_rf_all_abx2$n_abx_doses)] = 0
rownames(keep_for_rf_all_abx2) = keep_for_rf_all_abx2$sample.id; keep_for_rf_all_abx2$sample.id = NULL

keep_for_rf_each_abx2 = merge(keep_for_rf_each_abx,temp_df2, 
                            by = c("Child.ID","Date.of.collection"), all.y = T) %>%
  select(-Child.ID,-Date.of.collection,-symptomatic,-pre_detection)
keep_for_rf_each_abx2 = keep_for_rf_each_abx2[!is.na(keep_for_rf_each_abx2$sample.id),]
rownames(keep_for_rf_each_abx2) = keep_for_rf_each_abx2$sample.id; keep_for_rf_each_abx2$sample.id = NULL

abx_history = cbind(keep_for_rf_all_abx2,keep_for_rf_each_abx2)
rm(keep_for_rf_all_abx2,keep_for_rf_each_abx2)

#get metadata for only samples that have abx history
limited_metadata = as(sample_data(for_metadata)[,c("rainfall","drain","food_deficit",
                                                "drinking_water_source","income_mo",
                                                "water_treat","siblings_under_5",
                                                "month.of.birth","year.of.birth",
                                                "month.of.collection",
                                                "year.of.collection", "Any..Giardia",
                                                "Any..E.histolytica",
                                                "WAZ","WHZ","HAZ","BAZ",
                                                "Number.of.crypto.positive.events",
                                                "Age.in.days")],"data.frame")
limited_metadata$water_treat = as.factor(limited_metadata$water_treat)

```

# Can we predict whether a sample was sub-clinical or diarrheal from TOD microbiome AND ABx data?

```{r}

for_metadata = subset_samples(ps_rare_diff, Clinical.type %in% c(10,12))
for_metadata = subset_samples(for_metadata, !is.na(previous_surv))
temp_df = as(sample_data(for_metadata)[,c("sample.id")],"data.frame") 
rownames(temp_df) = temp_df$sample.id
temp_df2 = as(sample_data(for_metadata)[,c("Child.ID","Date.of.collection","sample.id")],"data.frame")
# kids with zero ABx don't have their Run# in these dataframes - cut those kids and then add after
keep_for_rf_all_abx2 = merge(keep_for_rf_all_abx %>% filter(!is.na(sample.id)), temp_df, 
                            by = 'sample.id', all.y = T) %>% select(sample.id,n_abx_doses)
keep_for_rf_all_abx2$n_abx_doses[is.na(keep_for_rf_all_abx2$n_abx_doses)] = 0
rownames(keep_for_rf_all_abx2) = keep_for_rf_all_abx2$sample.id; keep_for_rf_all_abx2$sample.id = NULL

keep_for_rf_each_abx2 = merge(keep_for_rf_each_abx,temp_df2, 
                            by = c("Child.ID","Date.of.collection"), all.y = T) %>%
  select(-Child.ID,-Date.of.collection,-symptomatic,-pre_detection)
keep_for_rf_each_abx2 = keep_for_rf_each_abx2[!is.na(keep_for_rf_each_abx2$sample.id),]
rownames(keep_for_rf_each_abx2) = keep_for_rf_each_abx2$sample.id; keep_for_rf_each_abx2$sample.id = NULL

abx_history = cbind(keep_for_rf_all_abx2,keep_for_rf_each_abx2)
rm(keep_for_rf_all_abx2,keep_for_rf_each_abx2)

x_withabx = cbind(cbind(abx_history, otu_table(for_metadata)), as(sample_data(for_metadata),"data.frame") %>% select(Age.in.days))
y = as.factor(sample_data(for_metadata)$Clinical.type)

rf = randomForest_res(x_withabx,y,25)
rf[[1]]

top20_TOD_with_abx = rf[[4]]
roc_TOD_abx = rf[[2]]
AUC_TOD_abx = rf[[3]]
AUC_TOD_abx

save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_TOD_abx_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

rm(rf,x_withabx,y)

```

## DATA NOT SHOWN: Can we predict whether a sample was sub-clinical or diarrheal from metadata AND ABx data?

```{r}

x = cbind(limited_metadata,abx_history)
y = as.factor(sample_data(for_metadata)$Clinical.type)

rf = randomForest_res(x,y,15)
rf[[1]]

top15_metadata_with_abx = rf[[4]]
roc_meta_abx = rf[[2]]
AUC_meta_abx = rf[[3]]
AUC_meta_abx

save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_meta_abx_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

rf = randomForest_res(limited_metadata,y,15)
rf[[1]]
top15_metadata_without_abx = rf[[4]]
roc_meta_no_abx = rf[[2]]
AUC_meta_no_abx = rf[[3]]
AUC_meta_no_abx

save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_meta_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

rm(rf,x,y,limited_metadata)

roc_list_mirpur = list(roc_meta_abx,roc_meta_no_abx)
names(roc_list_mirpur) = c("metadata","metadata &\nantibiotics")

```


# Prep for Figure 4D: Plot top 10 most important variables wth both microbiota classifiers

```{r}

## gini
plt_df1 = as.data.frame(top10_TOD) %>% rownames_to_column('variable') %>% dplyr::rename(value = top10_TOD) 
plt_df2 = as.data.frame(top10_PD) %>%rownames_to_column('variable') %>% dplyr::rename(value = top10_PD)
plt_df = rbind(plt_df1 %>% mutate(model = 'time-of-detection'),plt_df2 %>% mutate(model = 'pre-detection'))
plt_df$model = as.factor(plt_df$model)
save_df1 = plt_df1
save_df2 = plt_df2
save(list = c("save_df1","save_df2"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p2_" , format(Sys.time(), "%Y_%m_%d")),".RData")) 

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tidyr::spread(tax, key = model, value = value)  %>% # this line is new here
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))


plot_tax = tax2 %>% select(pre.detection, time.of.detection, var) %>% melt(id.vars = "var", na.rm = T)
plot_tax$var = as.factor(plot_tax$var)
plot_tax$var = reorder(plot_tax$var, -plot_tax$value)
plot_tax$var <- relevel(plot_tax$var, "ASV19 (Megasphaera)")
plot_tax_4D = plot_tax

```

## Save data for interaction plots - interested but DATA NOT SHOWN (Also, needs to be performed on a computer with more mem than my laptop)

```{r}

# save data for interaction plots
ASV = tax2[!(is.na(tax2$Genus))& tax2$Genus == "Megasphaera",1]
ASV_10 = tax2[,1]

tax = tax_table(ps_rare_diff) %>% as.data.frame()
tax$Genus = as.character(tax$Genus)
save_tax = tax %>% 
  rownames_to_column("Row.names") %>%
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')')))) 
save(list = c("ASV", "save_tax"), file = paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/megasphaera_ASV_"
              , format(Sys.time(), "%Y_%m_%d")
              , ".RData"))
save(list = c("ASV_10", "save_tax"), file = paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/top_ASVs_"
              , format(Sys.time(), "%Y_%m_%d")
              , ".RData"))

```

## DATA NOT SHOWN: Plot important variables (raw values)

```{r}

#extract ASV that are in Tax2 from predetection otu_table
otu_PD = as.data.frame(x_for_PD)[,plt_df2$variable] %>% rownames_to_column('sample') %>% mutate(symptomatic = y_for_PD)
#extract ASV that are in Tax2 from TOD otu_table
otu_TOD = as.data.frame(x_for_TOD)[,plt_df1$variable] %>% rownames_to_column('sample') %>%mutate(symptomatic = y_for_TOD)
ids = as(sample_data(ps_rare_diff),"data.frame") %>% select(Child.ID) %>% rownames_to_column('sample')
otu_PD = merge(otu_PD, ids, all.x = T, all.y = F, by = "sample")
otu_TOD = merge(otu_TOD, ids, all.x = T, all.y = F, by = "sample")

plt_df = melt(otu_TOD) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() %>%
  mutate(symptomatic_use = ifelse(symptomatic == '10','subclinical','diarrheal')) %>%
  tidyr::separate(sample, c("sample", "remove_extra")) %>%
  select(-remove_extra) 

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
#ASV10 (Collinsella) for PD
#ASV5 (Parabacteroides) for TOD
TOD_twoASVs = tax2 %>% filter(var == 'ASV10 (Megasphaera)' | var == 'ASV5 (Parabacteroides)')

tax2_TOD = tax2
rm(tax, tax2, plt_df)

plt_df = melt(otu_PD) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() %>%
  mutate(symptomatic_use = ifelse(symptomatic == '1','subclinical','diarrheal')) %>%
  tidyr::separate(sample, c("sample", "remove_extra")) %>%
  select(-remove_extra) 

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
#ASV10 (Collinsella) for PD
#ASV5 (Parabacteroides) for TOD
PD_twoASVs = tax2 %>% filter(var == 'ASV10 (Megasphaera)' | var == 'ASV3 (Collinsella)')
tax2_PD = tax2

rm(tax, tax2, plt_df)

```

# Prep for Figure 4E - important variables

```{r}

twoASVs = rbind(TOD_twoASVs %>% filter(var == "ASV10 (Megasphaera)") %>% mutate(time = 'time-of-detection'), 
                PD_twoASVs %>% filter(var == "ASV10 (Megasphaera)") %>% mutate(time = 'pre-detection'))

# trying to plot TOD on y axis and PD on x axis
temp1 = twoASVs %>% select(value_use,symptomatic_use, Genus,time, Child.ID) %>% filter(time == 'time-of-detection') %>% select(-time) %>% dplyr::rename(`time-of-detection` = value_use)
temp2 = twoASVs %>% select(value_use,symptomatic_use, Genus,time, Child.ID) %>% filter(time == 'pre-detection') %>% select(-time) %>% dplyr::rename(`pre-detection` = value_use)
two_WIDE = merge(temp1, temp2, all = T)
colnames(two_WIDE)[colnames(two_WIDE) == 'symptomatic_use'] = 'symptomatic?'

twoASVs_TOD_wide = tidyr::spread(TOD_twoASVs %>% select(value_use,symptomatic_use,sample, Genus),
                             key = Genus, value = value_use) %>% 
  mutate(`symptomatic?` = as.factor(symptomatic_use)) 

twoASVs_PD_wide = tidyr::spread(PD_twoASVs %>% select(value_use,symptomatic_use,sample, Genus),
                             key = Genus, value = value_use) %>% 
  mutate(`symptomatic?` = as.factor(symptomatic_use))

count_df_PD_mirpur = twoASVs_PD_wide
count_df_TOD_mirpur = twoASVs_TOD_wide
save(list = c("count_df_PD_mirpur","count_df_TOD_mirpur"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p3_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

```

## DATA NOT SHOWN: Imp var for first v later infection

```{r}
 
plt_df1 = as.data.frame(topN_first_infection) %>% rownames_to_column('variable') %>% dplyr::rename(value = topN_first_infection) 
plt_df2 = as.data.frame(topN_first_infection_PD) %>%rownames_to_column('variable') %>% dplyr::rename(value = topN_first_infection_PD)
plt_df = rbind(plt_df1 %>% mutate(model = 'time-of-detection'),plt_df2 %>% mutate(model = 'pre-detection'))
plt_df$model = as.factor(plt_df$model)

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))

tax2_infect = tax2
rm(tax, tax2, plt_df)

```

# Prep Supplemental Figure 5 - ROC and Important variables, Adding in Abx to model

```{r}

plt_df1 = as.data.frame(top10_TOD) %>%rownames_to_column('variable') %>% dplyr::rename(value = top10_TOD) 
plt_df2 = as.data.frame(top20_TOD_with_abx) %>%rownames_to_column('variable') %>% dplyr::rename(value = top20_TOD_with_abx) %>% top_n(10,value)
plt_df = rbind(plt_df1 %>% mutate(model = 'without abx'),plt_df2 %>% mutate(model = 'with abx'))
plt_df$model = as.factor(plt_df$model)

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
tax2_abx = tax2
rm(tax,tax2,plt_df,plt_df1,plt_df2)

```

## DATA NOT SHOWN: Can we predict whether a sample was sub-clinical or diarrheal from microbiota based on the change in microbiome?

```{r}

compare <- c(10,12) # TOD +/- symptoms
ps_for_class = change_in_microbiota_ps
ps_for_class <- subset_samples(ps_for_class, Clinical.type %in% compare)
ps_for_class_rel = ps_for_class#transform_sample_counts(ps_for_class, function(x) x/sum(x))

x <- otu_table(ps_for_class_rel)
y <- as.factor(sample_data(subset_samples(ps_for_class_rel, Clinical.type %in% compare))$Clinical.type)


rf = randomForest_res(x,y,10)
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_change_" , format(Sys.time(), "%Y_%m_%d")),".RData"))     
roc_change = rf

rm(x,y,rf,ps_for_class_rel,ps_for_class,p)

```

# Save data for other scripts

```{r}

#"roc_list_mirpur","otu_TOD","","",
# 
# save(list = c("predetection_tax2","tod_tax2"), file = paste0(paste0("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

save(list = c("chi_df","roc_TOD","roc_PD","plot_tax_4D","otu_PD","tax2_TOD","tax2_PD","twoASVs","two_WIDE","tax2_infect","roc_TOD","roc_TOD_abx","tax2_abx","roc_change","roc_allM_M","roc_allM","AUC_TOD","AUC_PD","AUC_allM","AUC_allM_M","AUC_TOD_abx"), file = paste0(paste0("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

```

# Make figures
```{r}

#rmarkdown::render("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/code/R_markdown/P3_machine_learning_figures.Rmd")

```
