---
title: "Microbiome analysis - machine learning"
author: "Maureen A. Carey"
date: "09/01/2020"
output: html_document
---

# Load packages and data

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/")

# load packages
library(ggplot2); library(pROC); library(dplyr); library(tibble); library(reshape2)

# set constants
SEED_INT = 10; set.seed(SEED_INT); options(scipen=10000); theme_set(theme_bw())
colors = c("#1F78B4","#FF7F00","#666666", 'light gray')
names(colors) = c('diarrheal','subclinical','no crypto','both')
# blue = "#1F78B4" orange = "#FF7F00" gray = "#666666"

# load data
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_2020_09_02.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_2020_09_10.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/cmel_P3output_plot_2020_09_11.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/chom_P3output_plot_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_2020_09_11.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/WOC_P3output_plot_2020_09_11.RData")

```

# Figure 4C: ROC Curve

```{r}

roc_list = list(roc_TOD,roc_PD)
names(roc_list) = c("TOD microbiota","pre-detection microbiota")
p = pROC::ggroc(roc_list, aes = "colour") + 
  theme_bw() +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed") +
  scale_color_manual(values = c("blue","red"), guide = F) + 
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.title = element_blank()) +
  annotate("text", label = paste0("TOD AUC = ",round(AUC_TOD, digits = 2)), x = .25, y = .27, color = "blue") +
  annotate("text", label = paste0("PD AUC = ",round(AUC_PD, digits = 2)), x = .25, y = .13, color = "red") 
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4C_ROC_microbiome_mirpur.png",
       p,width = 3.2,height = 3,units = "in",dpi = 300)

```

# Figure 4D: Plot top 10 most important variables wth both microbiota classifiers

```{r}

plot_tax_4D %>%
  ggplot() +geom_bar(aes(x=var, y = value, fill = variable),stat = "identity", width=0.4, position = position_dodge(width=0.5)) +
  scale_fill_manual(values = c("red","blue")) +
  xlab(NULL) + ylab('Mean decrease in node impurity') + #facet_wrap(~tod_or_pre, scales = "free_x") +
  coord_flip() + 
  theme(legend.position = c(.76, 0.13), 
        axis.text.x = element_text(size = 7),
        legend.title=element_blank(), 
        legend.text=element_text(size=8))  + scale_x_discrete(limits = rev(levels(plot_tax_4D$var)))
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4D_imp_var_microbiome_mirpur.png",
       width = 4.5,height = 4,units = "in",dpi = 300)

```

## DATA NOT SHOWN: Plot important variables (raw values)

```{r}

# #plot PD
# reshape2::melt(otu_PD %>% rownames_to_column('sample')) %>%
#   group_by(variable) %>%
#   mutate(value_use = value/max(value)) %>%
#   ungroup() %>%
#   mutate(symptomatic_use = ifelse(symptomatic == '10','subclinical','diarrheal')) %>%
#   ggplot(aes(x = reorder(variable,value), y = value_use, group = sample, color = symptomatic_use)) +
#   geom_point(alpha = .5) + geom_line(alpha = .5) +
#   scale_color_manual(values = c(colors['diarrheal'],colors['subclinical'])) +
#   ylab('relative abundance') + xlab(NULL) + labs(color = "") + scale_y_sqrt()

#plot TOD
ggplot(tax2_TOD, aes(x = reorder(var,value), y = value_use, color = symptomatic_use)) +
  geom_boxplot(position = "dodge", outlier.shape = NA) +
  geom_point(alpha = .5, size = 1,
             position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0, 
                                             seed = SEED_INT, dodge.width = 0.75)) +
  scale_color_manual(values = c(colors['diarrheal'],colors['subclinical'])) +
  ylab('relative abundance') + xlab(NULL) + labs(color = "")+ scale_y_sqrt() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust = 0))
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/TOD_imp_var_boxplot_mirpur.png",
       width = 5,height = 3,units = "in",dpi = 300)

#plot PD
ggplot(tax2_PD, aes(x = reorder(var,value), y = value_use, color = symptomatic_use)) +
  geom_boxplot(position = "dodge", outlier.shape = NA) +
  geom_point(alpha = .5, size = 1,
             position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0, 
                                             seed = SEED_INT, dodge.width = 0.75)) +
  scale_color_manual(values = c(colors['diarrheal'],colors['subclinical'])) +
  ylab('relative abundance') + xlab(NULL) + labs(color = "")+ scale_y_sqrt() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0, vjust = 0))
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/PD_imp_var_boxplot_mirpur.png",
       width = 5,height = 3,units = "in",dpi = 300)

```

# Figure 4E - important variables

```{r}

ggplot(twoASVs, aes(x = var, y = value_use, color = symptomatic_use)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.5, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0, dodge.width = 0.75, seed = SEED_INT)) + 
  facet_wrap(~time, scales = "free_x") +
  scale_color_manual(values= colors[1:2]) +
  scale_y_sqrt() + ylab('relative abundance') + xlab(NULL) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        legend.position = "bottom",
        legend.title = element_text(size = 9)) + guides(color=guide_legend(title="infection")) +
  geom_hline(yintercept = 0.2, linetype = 'dotted', alpha = .75)
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4E_mega.png",
       width = 3.2,height = 4,units = "in",dpi = 300)

```

## DATA NOT SHOWN: Imp var for first v later infection

```{r}

data_wide = tidyr::spread(chi_df, symptomatic, Freq)
rownames(data_wide) = data_wide$n_crypto; data_wide$n_crypto = NULL
res = chisq.test(data_wide) 
chi_df %>% ggplot(aes(x = n_crypto, y = Freq, fill = symptomatic)) + geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = colors) + ylab('episodes') + xlab('first or subsequent infection?') +
  annotate(geom = "text", x = 2, y = 45, label = paste0("Pearson's Chi-\nsquared test\np-value = ",round(res$p.value,3))) + ylim(0,52) + theme(legend.position = "bottom")

tax2_infect %>%
  ggplot() +geom_bar(aes(x=reorder(var,value), y = value, fill = model),stat = "identity", width=0.4, position = position_dodge(width=0.5)) +
  scale_fill_manual(values = c("red","blue")) +
  xlab(NULL) + ylab('Mean decrease in node impurity') + #facet_wrap(~tod_or_pre, scales = "free_x") +
  coord_flip()
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/imp_var_Infect#_mirpur.png",
              width = 7,height = 3,units = "in",dpi = 300)

```

## DATA NOT SHOWN: metadata doesn't help classifier

```{r}

roc_list = list(roc_allM_M,roc_allM)
names(roc_list) = c("all microbiota & metadata","all microbiota")
p = pROC::ggroc(roc_list, aes = "colour") +
  theme_bw() +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed") +
  scale_color_manual(values = c("grey","red")) +
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.title = element_blank()) +
  annotate("text", label = paste0("AUC = ",round(AUC_allM, digits = 2)), x = .2, y = .13, color = "red") +
  annotate("text", label = paste0("AUC = ",round(AUC_allM_M, digits = 2)), x = .2, y = .2, color = "grey")
p

```

# Supplemental Figure 5 - ROC and Important variables, Adding in Abx to model

```{r}

load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_TOD_abx_2020_09_11.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/WOC_rf_TOD_abx_2020_09_11.RData")

AUC_TOD_abx = rf[[3]]
rm(rf)
roc_list = list(roc_TOD,roc_TOD_abx)
names(roc_list) = c("TOD microbiota","TOD microbiota &\nantibiotic history")
p = pROC::ggroc(roc_list, aes = "colour") + 
  theme_bw() +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed") +
  scale_color_manual(values = c("blue","grey")) + 
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.title = element_blank()) +
  annotate("text", label = paste0("AUC = ",round(AUC_TOD, digits = 2)), x = .2, y = .27, color = "blue") +
  annotate("text", label = paste0("AUC = ",round(AUC_TOD_abx, digits = 2)), x = .2, y = .2, color = "grey") 
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig5A_ROC_TOD_abx_mirpur.png",
       p,width = 5,height = 3,units = "in",dpi = 300)

tax2_abx %>%
  ggplot() +geom_bar(aes(x=reorder(var,value), y = value, fill = model),stat = "identity", width=0.4, position = position_dodge(width=0.5)) +
  scale_fill_manual(values = c("grey","blue")) +
  xlab(NULL) + ylab('Mean decrease in node impurity') + #facet_wrap(~tod_or_pre, scales = "free_x") +
  coord_flip()
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig5B_imp_var_abx_mirpur.png",
       width = 7,height = 3,units = "in",dpi = 300)

```

## DATA NOT SHOWN: Can we predict whether a sample was sub-clinical or diarrheal from microbiota based on the change in microbiome?

```{r}

p = pROC::ggroc(roc_change[[2]]) + 
  theme_bw() +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed") + 
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm")) + guides(color = F) + ggtitle(roc_change[[3]])
p


```
