---
title: "Response to Reviewers - Part 2 (Higher Taxonomic level)"
author: "Maureen A. Carey"
date: "1/14/2021"
output: html_document
---

# Load packages and data

```{r setup, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/")

# load packages
library(phyloseq); library(reshape2); library(tibble)
library(randomForest); library(rmarkdown); library(dplyr); library(tidyr); library(pROC)

# set constants
SEED_INT = 10; set.seed(SEED_INT); options(scipen=10000)
colors = c("#1F78B4","#FF7F00","#666666", 'light gray')
names(colors) = c('diarrheal','subclinical','no crypto','both')
# blue = "#1F78B4" orange = "#FF7F00" gray = "#666666"

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
total_diarrhea = total_diarrhea  %>% filter(CHILDID %in% burden_mirpur$SID) # remaining kids never were sick
all_cause_diar = total_diarrhea %>% group_by(CHILDID,study) %>% summarize(episodes = n()) %>% ungroup()
rm(list= ls()[! (ls() %in% c('all_cause_diar','SEED_INT','colors'))])
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P2output_2020_09_11.RData")

# make useful functions
randomForest_res = function(x,y,N) { 
  # x = variables used to predict class, y = classs labels, N = number imp variables you want
  # THIS ONLY WORKS IF YOU"RE CLASSIFYING 10 v. 12 (subclinical v. diarrheal)
  rf_res_object = randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.0001,100))

  extracted_importance = importance(rf_res_object,type=2) # type 2 = node impurity, more useful for auc
  extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
  topN = tail(extracted_importance,N)

  predictions = as.data.frame(predict(rf_res_object,type="prob"))
  predictions$predict = names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
  predictions$observed = y
  
  ROC_res = roc(ifelse(predictions$observed==10, 10, 12), as.numeric(predictions$`10`))
  AUC_res = auc(ROC_res)
  return(list(rf_res_object,ROC_res,AUC_res,topN))  
}

```

# Higher taxonomy

```{r}

ps_rare_diff = tax_glom(ps_rare_diff, "Genus")
#ps_rare_diff = tax_glom(ps_rare_diff, "Family")

```

# Can we predict whether a sample was sub-clinical or diarrheal from predetection microbiota?

```{r}

compare <- c(1,2) # predetection -> +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)

x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
rf <- randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.00015,130))

save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/genus_level_rf_PD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))  

x_for_PD = x
y_for_PD = y

extracted_importance = importance(rf,type=2) # type 2 = node impurity, more useful for auc
extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
top10_PD = tail(extracted_importance,15)

predictions = as.data.frame(predict(rf,type="prob"))
predictions$predict <- names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
predictions$observed <- y

roc_PD <- roc(ifelse(predictions$observed==1, 1, 2), as.numeric(predictions$`1`))
AUC_PD = auc(roc_PD)
AUC_PD

rm(x,y,rf,predictions,extracted_importance,ps_for_class)

```

# Can we predict whether a sample was sub-clinical or diarrheal from TOD microbiota?

```{r}

compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
x_for_TOD = x
y_for_TOD = y

rf = randomForest_res(x_for_TOD,y_for_TOD,15)
top10_TOD = rf[[4]]
roc_TOD = rf[[2]]
AUC_TOD = rf[[3]]
AUC_TOD

save(list = c("rf","x","y"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/genus_level_rf_TOD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))
rm(x,y,rf,ps_for_class)

```

# Is Megasphaera an important variable with this genus level analysis?

```{r}

## gini
plt_df1 = as.data.frame(top10_TOD) %>% rownames_to_column('variable') %>% dplyr::rename(value = top10_TOD) 
plt_df2 = as.data.frame(top10_PD) %>%rownames_to_column('variable') %>% dplyr::rename(value = top10_PD)
plt_df = rbind(plt_df1 %>% mutate(model = 'time-of-detection'),plt_df2 %>% mutate(model = 'pre-detection'))
plt_df$model = as.factor(plt_df$model)

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tidyr::spread(tax, key = model, value = value)  %>% # this line is new here
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))


important_genera = na.omit(tax2)
plot_tax = tax2 %>% select(pre.detection, time.of.detection, var) %>% melt(id.vars = "var", na.rm = T)
plot_tax$var
grep("Megasphaera", levels(plot_tax$var))

```

# No (see megasphaera prevalence plots in response_to_reviewers_p1.Rmd)

Rather, `r important_genera$Genus` are important at both timepoints. Many of these genera were  important in the ASV-level time-of-detection classifer, but some emerge as important only when looking at the genus-level analysis. None of these genera were important in the ASV-level predetection classifier and only Sutterella, Bifidobacterium, and Veillonella were important in the ASV-level time-of-detection classifier.